Gravitational Lens Modelling of Flux Ratio Anomalous System B1555+375

### 2015-04-23 papers/B1555_EdgeOn/B1555_letter.tex
Update of Simona's change

Gravitational lens flux-ratio anomalies provide a powerful technique
for detecting the presence of dark matter substructure in distant
galaxies.  However, before using the statistics of systems with these
anomalies, it is imperative to ascertain that a given system does not
have properties that could plausibly cause the observed anomaly
without the need for substructure.  In this letter, we present the
case of the B1555+375 lens system, which has a strong radio-wavelength
flux-ratio anomaly but where the lensing galaxy is also revealed in
high-resolution imaging to have a clear edge-on disk component.  The
disk crosses directly over the pair of images that exhibit the
anomaly, and we show that simple models that include the disk can
reproduce the anomaly without requiring substructure.  Several [??]
other lens systems that we have observed with adaptive optics imaging
as part of the Strong lensing at High Angular Resolution Programme (SHARP)
show similar behaviour.  These results suggest that the assumption that
all flux-ratio anomalies are due to substructure will likely overestimate
the substructure fraction in massive galaxies.

@@

It was first suggested by \citet{Mao1998} that flux ratio anomalies observed in radio loud multiple images of lensed quasar could be interpreted as the telltale of the presence of substructure in lens galaxies.
Indeed, small perturbations in the gravitational potential are
sufficient to bring flux ratio anomalies in strong lensed systems \citep{Bradac02}. 
Moreover, unlike in the optical or infrared, in the radio band, lensed
images are not sensitive to micro lensing from stars nor dust
extinction. This makes flux ratio anomalies a promising tool to
detect dark matter substructures \citep{Dalal2002, N13}.  At present,
the amount of substructure derived from flux ratio anomalies is in
disagreement with predictions from numerical simulations, with the
former requiring a larger fraction of mass in substructure than
expected \citet{Xu14}. While this could be the result of small sample
statistics, it could also be the indication that flux ratio anomalies
have a different origin than substructure. 
%
An alternative method for
the detection of substructure in gravitational lens galaxies based on
the surface brightness distribution of extended arcs and Einstein rings
has been introduced by \citet{K05} and \citet{V09}. Interestingly, the fraction of
substructure measured with this \emph{gravitational imaging technique}
is currently in agreement with theoretical expectations \citep{V14a}.

@@

Gravitational lensing is, {\bf at present,} the most powerful tool for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes{\bf, such as propagation effects and assumptions on the mass distribution of the lens galaxy}.  
It is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal
\citep[although see][for a rare example of radio microlensing]{K2000}.
Radio propagation effects can affect radio fluxes, as is clearly shown
in the B0128+437 system \citep{B04}.  However, these propagation
effects can be disentangled from substructure lensing because they
have well known dependencies on wavelength. \citet{KD04} examine all
eight radio loud flux anomalous lenses besides B0128+437, namely: MG
0414+0534, B0712+472, B1359+154, B1422+231, B1555+375, B1608+656,
B1933+503, and B2045+265.  They find no strong wavelength dependence
in the flux ratios of all these systems and therefore rule out radio
propagation effects as significant contributors to the observed
anomalies.

@@

From a theoretical perspective, \citet{Xu14} find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. By adding subhaloes from numerical simulations
on to these systems, only one out of eight radio loud systems
reproduces the level of flux anomalies seen in the data.  The rest of
them are unlikely (only $1 \% - 4 \%$ probability) to have flux
anomalies produced by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} conclude the failure in
predicting flux ratios is due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in lens model. Their results are consistent with the SHARP
$K^{\prime}$ band observations, in which an edge-on disk crosses the
radio merging double.

Additional near-infrared AO imaging from SHARP shows that B1555+375 is
not the only system having a significant disk component detection in
the lensing galaxy.  These systems will be investigated in
future work, but the images alone strongly suggest that other lens
systems have similar links between flux-ratio anomalies and disk
components.  Although the number of radio loud flux anomalous lensed
quasars remains small, a new technique using narrow-line emission
regions will enlarge the sample and bring statistical perspective on
substructures in the near future \citep{N14}. It is well known (e.g,
\citep{Ka91}) that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, derived
substructure fractions should be considered as upper limits.

\section{Conclusions}
{\bf In this letter we have taken advantage of the power of multi-wavelength high resolution data to investigate whether flux ratio anomalies could be explained by more complex smooth models for the lens galaxy mass distribution rather than substructure. Thanks to new imaging from the SHARP $K^{\prime}$ band observations of the gravitational lens system B1555+375, we have identified the existence of a previously undetected edge-on disk. By including this disk in the mass model of the lens galaxy, we find a good fit for both the positions and flux ratios of the lensed images. Our results clearly demonstrate that the lack of knowledge on the lens galaxy mass distribution, combined with the limited information provided by the position of a few lensed images, can lead to improper lens models and a misinterpretation of the origin of flux ratio anomalies. 
We stress, therefore, that one should be very careful when interpreting the nature of flux ratio anomalies, as prior assumptions on the lens mass distribution are a significant source of bias. Finally, our analysis shows that in order to derive precise quantification of the substructure population and any derived properties from strongly lensed flux anomalous systems, it is crucial to have a careful estimation of non-substructure effects. In fact, even in the presence of mass substructures, the observed flux ratio anomalies could be still dominated by the complex structure of the lens galaxy.}

>>>>

Gravitational lens flux-ratio anomalies provide a powerful technique
for detecting the presence of dark matter substructure in distant
galaxies.  However, before using the statistics of systems with these
anomalies, it is imperative to ascertain that a given system does not
have properties that could plausibly cause the observed anomaly
without the need for substructure.  In this letter, we present the
case of the B1555+375 lens system, which has a strong radio-wavelength
flux-ratio anomaly but where the lensing galaxy is also revealed in
high-resolution imaging to have a clear edge-on disk component.  The
disk crosses directly over the pair of images that exhibit the
anomaly, and we show that simple models that include the disk can
reproduce the anomaly without requiring substructure.  
Other lens systems that we have observed with adaptive optics imaging
as part of the Strong lensing at High Angular Resolution Programme (SHARP)
show a similar behaviour and will be the subject of a future paper.  
These results suggest that the assumption that
all flux-ratio anomalies are due to substructure will likely overestimate
the substructure fraction in massive galaxies.

@@

It was first suggested by \citet{Mao1998} that flux ratio anomalies observed in radio loud multiple images of lensed quasar could be interpreted as the telltale of the presence of substructure in lens galaxies.
Indeed, small perturbations in the gravitational potential are
sufficient to bring flux ratio anomalies in strong lensed systems \citep{Bradac02}. 
Moreover, unlike in the optical or infrared, in the radio band, lensed
images are not sensitive to micro lensing from stars nor dust
extinction. This makes flux ratio anomalies a promising tool to
detect dark matter substructures \citep{Dalal2002, N13}.  At present,
the amount of substructure derived from flux ratio anomalies is in
disagreement with predictions from numerical simulations, with the
former requiring a larger fraction of mass in substructure than
expected \citep{Xu14}. While this could be the result of small sample
statistics, it could also be the indication that flux ratio anomalies
have a different origin than substructure. 
%
An alternative method for
the detection of substructure in gravitational lens galaxies based on
the surface brightness distribution of extended arcs and Einstein rings
has been introduced by \citet{K05} and \citet{V09}. Interestingly, the fraction of
substructure measured with this \emph{gravitational imaging technique}
is currently in agreement with theoretical expectations \citep{V14a}.

@@

Gravitational lensing is, {\bf at present,} the most powerful tool for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes{\bf, such as propagation effects and assumptions on the mass distribution of the lens galaxy}.  
It is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal
\citep[although see][for a rare example of radio microlensing]{K2000}.
Radio propagation effects can affect radio fluxes, as is clearly shown
in the B0128+437 system \citep{B04}.  However, these propagation
effects can be disentangled from substructure lensing because they
have well known dependencies on wavelength. \citet{KD04} examine all
eight radio loud flux anomalous lenses besides B0128+437, namely: MG
0414+0534, B0712+472, B1359+154, B1422+231, B1555+375, B1608+656,
B1933+503, and B2045+265.  They found no strong wavelength dependence
in the flux ratios of all these systems and therefore rule out radio
propagation effects as significant contributors to the observed
anomalies.

@@

From a theoretical perspective, \citet{Xu14} find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. By adding subhaloes from numerical simulations
on to these systems, only one out of eight radio loud systems
reproduces the level of flux anomalies seen in the data.  The rest of
them are unlikely (only $1 \% - 4 \%$ probability) to have flux
anomalies produced by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} have concluded that the failure in
predicting flux ratios is most likely due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in the lens model. Their results are consistent with the SHARP
$K^{\prime}$ band observations, in which an edge-on disk crosses the
radio merging double and which fully accounts for the observed flux ratio anomaly.  In general, it is well known (e.g,
\citep{Ka91}) that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, derived
substructure fractions should be considered as upper limits.

Although the number of radio loud flux anomalous lensed
quasars remains small, a new technique using narrow-line emission
regions will enlarge the sample and bring statistical perspective on
substructures in the near future \citep{N14}.

\section{Conclusions}
{\bf In this letter we have taken advantage of the power of multi-wavelength high resolution data to investigate whether flux ratio anomalies could be explained by more complex smooth models for the lens galaxy mass distribution rather than substructure. Thanks to new imaging from the SHARP $K^{\prime}$ band observations of the gravitational lens system B1555+375, we have identified the existence of a previously undetected edge-on disk. By including this disk in the mass model of the lens galaxy, we find a good fit for both the positions and flux ratios of the lensed images. Our results clearly demonstrate that the lack of knowledge on the lens galaxy mass distribution, combined with the limited information provided by the position of a few lensed images, can lead to improper lens models and a misinterpretation of the origin of flux ratio anomalies. 
We stress, therefore, that one should be very careful when interpreting the nature of flux ratio anomalies, as prior assumptions on the lens mass distribution are a significant source of bias. Our analysis shows that in order to derive precise quantification of the substructure population and any derived properties from strongly lensed flux anomalous systems, it is crucial to have a careful estimation of non-substructure effects. In fact, even in the presence of mass substructures, the observed flux ratio anomalies could be still dominated by the complex structure of the lens galaxy. Finally, additional near-infrared AO imaging from SHARP shows that B1555+375 is not the only system having a significant disk component detection in
the lensing galaxy.  These systems will be investigated in future work, but the images alone strongly suggest that other lens systems have similar links between flux-ratio anomalies and disk components. 
}
